---
title: "Shiny Dashboards"
subtitle: "RSS 2020 - Visualising Data with Dashboards"
author: "Dr. Aidan Boland"
date: "`r format.Date('2020-09-07', '%d %B %Y')`"
output:
  ioslides_presentation:
    css: files/mystyle.css
    fig_width: 10
    # logo: files/Edge_Logo_flat.png
    smaller: no
    widescreen: yes
# runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.align  = 'center', dev.args = list(bg = 'transparent'), cache = F)
library(png)
library(grid)
```


##

<br><br>

- Introduction
<br><br><br>

- Example of Shiny Apps
<br><br><br>

- Deployment



# Introduction


## About Me {.columns-2 .smaller}

```{r , fig.height=1, fig.width=5, fig.align='left'}
img <- readPNG("files/UCD_logo.png")
grid.raster(img)
```
<br><br>
```{r insight centre logo, fig.height=1, fig.width = 5, fig.align='left'}
img <- readPNG("files/insight-centre.png")
grid.raster(img)
```
<br><br>
```{r clavis logo, fig.height=1, fig.width = 4, fig.align='center'}
img <- readPNG("files/Clavis-Insight.png")
grid.raster(img)
```

```{r , fig.height=1, fig.width=3, fig.align='center'}
img <- readPNG("files/Edge_Logo.png")
grid.raster(img)
```




- BSc (Mathematical Sciences)
- MSc (Statistics)
- PhD (Computational Statistics)
    - Prof. Nial Friel
<br><br>
- PostDoc
   - Insight Centre for Data Analytics (UCD)
   - Prof. Andrew Parnell
<br><br>
- Data Scientist (Clavis Insight)
   - Ecommerce analytics startup
   - Acquired by Ascential 2017
<br><br>
- Senior Data Scientist (Edge by Ascential)
   - Combination of 4 companies


## My Journey with Shiny

- First created a Shiny App in 2015
  - [Teaching app](http://54.170.29.71:3838/TeachingApp/) to help explain regression and correlation




## {.shinyslide}

```{r teaching url, screenshot.opts = list(delay = 5), cache = F}
knitr::include_app("https://aboland.shinyapps.io/teaching_app/", height = "800px")
```


<!-- ##  {.shinyslide} -->

<!-- ```{r, echo = FALSE, cache = F, dev.args = list(bg = 'transparent')} -->
<!-- shinyAppDir( -->
<!--   "teaching_app/", -->
<!--   options = list(width = "100%", height = 7) -->
<!-- ) -->
<!-- ``` -->



## My Journey with Shiny

- First created a Shiny App in 2015
  - [Teaching app](https://aboland.shinyapps.io/teaching_app/) to help explain regression and correlation

- Second App created during Postdoc 
  - Needed a way of allowing a non technical user to run my code
  


## {.imageslide}

```{r autoc shiny image, fig.height=6}
img <- readPNG("files/shiny_example1.png")
grid.raster(img)
```


## My Journey with Shiny

- First created a Shiny App in 2015
  - Teaching app to help explain regression and correlation

- Second App created during Postdoc 
  - Needed a way of allowing a non technical user to run my code

- Current Role
  - 6 Shiny apps in use internally
  - Various uses
     - Run ML algorithms
     - ETL workflows
     - Database access


## My Journey with Shiny

- First created a Shiny App in 2015
  - Teaching app to help explain regression and correlation

- Second App created during Postdoc 
  - Needed a way of allowing a non technical user to run my code

- Current Role
  - 6 Shiny apps in use internally
  - Various uses
     - Run ML algorithms
     - ETL workflows
     - Database access
- Personal Projects
  - Exploring and visualising data




# Dashboards   

## Examples

- Football Data
   - [github.com/aboland/footballR](https://github.com/aboland/footballR)
   - Premier League Data for last 20 years
   - Various Statistics
<br><br>
- Dublin Bus
  - [github.com/aboland/DublinRTPI](https://github.com/aboland/DublinRTPI)
  - Uses Public API to get real time bus/train info
  - Collate info for multiple stops and routes


## {.shinyslide}

```{r pl app, screenshot.opts = list(delay = 5), cache = F}
knitr::include_app("https://aboland.shinyapps.io/football/", height = "800px")
```


## [Plotly](https://plotly.com/r/)

- Graphing library
  - R, Python, JavaScript
- Has integration with [ggplot2](https://ggplot2.tidyverse.org/) and [shiny](https://shiny.rstudio.com/)
- Plots can be also be saved as HTML files or displayed within R Markdown HTML documents or presentations


## {.imageslide}

<!-- labs(x = "engine displacement", y = "highway mpg", colour = "car type") -->
```{r ggplot example, echo = T, warning=FALSE, dev.args = list(bg = 'transparent')}
library(ggplot2)
my_plot <- 
  ggplot(mpg, aes(x = displ, y = hwy, colour = class)) +
  geom_point()
my_plot
```


## {.imageslide}


```{r plotly example, echo = T, warning=FALSE, message=FALSE}
library(plotly)
my_plot <- 
  ggplot(mpg, aes(x = displ, y = hwy, colour = class)) +
  geom_point()
ggplotly(my_plot)
```


## {.imageslide}

```{r plotly example2, echo = T, warning=FALSE, message=FALSE}
library(plotly)
my_plot <- 
  ggplot(mpg, aes(x = displ, y = hwy, colour = class)) +
  geom_point() + theme(plot.background = element_rect(fill = "transparent", color = NA))
ggplotly(my_plot)
```


## app.R {.imageslide}


```{r eval = FALSE, echo = TRUE}
library(shiny)
library(plotly)

ui <- fluidPage(
    plotlyOutput("sample_plot")
)

server <- function(input, output) {
    output$sample_plot <- renderPlotly({
        my_plot <- 
            ggplot(mpg, aes(displ, hwy, colour = class)) +
            geom_point()
        ggplotly(my_plot)
    })
}

# Run the application 
shinyApp(ui = ui, server = server)
```


## [Plotly](https://plotly.com/r/) - Save as HTML

<br><br>

```{r plotly example save, echo = T, eval = F}
library(htmlwidgets)
saveWidget(ggplotly(my_plot), "~/Desktop/sample_plotly.html")
```



## Dublin RTPI

- Real Time Passenger Information
- Entry for RStudio's Shiny Contest 2019
- Obtains data from public API
- Displays data for multiple bus stops



## {.imageslide}

```{r db official, fig.height=6}
img <- readPNG("files/db_official.png")
grid.raster(img)
```


## {.shinyslide}

```{r db app, screenshot.opts = list(delay = 5), cache = F}
knitr::include_app("https://aboland.shinyapps.io/DublinTransport/", height = "800px")
```


## URL Header Variables

- [aboland.shinyapps.io/DublinTransport/](https://aboland.shinyapps.io/DublinTransport/)
- [aboland.shinyapps.io/DublinTransport/?stops=334,336&routes=14,140](https://aboland.shinyapps.io/DublinTransport/?stops=334,336&routes=14,140)

<br><br>

```{r sample header code, eval = F, echo = T}
...
query <- parseQueryString(session$clientData$url_search)

if ("routes" %in% names(query)) {
  selected_stops <- unlist(strsplit(query$stops, ","))
} else{
  selected_stops <- c(334, 336)  # Use as default
}
...
```




# Deployment




## Deployment

- [shinyapps.io](https://www.shinyapps.io/) 
- [Amazon Web Services](https://aws.amazon.com/)


## [shinyapps.io](https://www.shinyapps.io/)   {.columns-2}

- **Free**
  - 5 applications
  - 12 active hours
- **Starter** $100/yr $\approx$ £75
  - 25 applications
  - 100 active hours
- **Basic** $440/yr $\approx$ £333
  - Unlimited applications
  - 500 active hours
    
<br><br>

- Pros
  - Simple to deploy app
  - Almost no maintenance
  - Easy to start and stop apps
  - Visibility on app usage
- Cons
  - Non customisable URL's  
  (without pro plan)
  - App limits
  - Usage limits

## {.imageslide}

```{r shiny_dep_img, fig.height=6}
img <- readPNG("files/shinyio_dep.png")
grid.raster(img)
```


## {.imageslide}

```{r shiny_dash_img, fig.height=6}
img <- readPNG("files/shinyio_dash.png")
grid.raster(img)
```


## Amazon Web Services  {.columns-2}

<br><br><br><br>

- Server cost
   - $94.26  
   (EC2 t3.micro + 8GB storage)
   - €33.76 [aboland.ie](http://aboland.ie/) domain
   - Total €114 $\approx$ £102
   
<br><br><br><br>

- Pros
   - Flexibility
   - Custom URL's
   - Can use server for other uses e.g. website, storage, API's
- Cons
   - Harder to deploy apps
   - Server maintenance
      - AWS management
      - Machine libraries
   - Harder to get visibility on app usage


## Thank you

<br><br>

- `r icon::fa("linkedin")` [ linkedin.com/in/aidanboland](https://www.linkedin.com/in/aidanboland)  
- `r icon::fa("twitter")` [ \@BolandAi](https://twitter.com/BolandAi)  
- `r icon::fa("github")` [ github.com/aboland](https://github.com/aboland)  
- `r icon::fa("link")` [ aboland.ie](http://aboland.ie/)

<br>

- `r icon::fa("link")` [github.com/aboland/RSS2020](https://github.com/aboland/RSS2020)
   - presentation.Rmd








  
